Hilbert space methods to approximate Gaussian processes using Stan

### 2019-09-07 Paper/manuscript/manuscript_05.tex

\usepackage{subfigure}
\usepackage{multirow}
\usepackage{float}
\usepackage{soul}
\usepackage{xcolor}
\graphicspath{{./images/}}

@@

Ultimately, these recommendations lie on the relationships among the number of basis functions, the boundary factor and the lengthscale of the function, which depend on the particular choice of the kernel function. In this work we built these relationships for the square exponential covariance function. For other kernels, the relationships will be slightly different.

@@

\begin{figure}
\begin{center}
\begin{tabular}{ c c c}
\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part1.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part1.pdf} & \hspace{-6mm} \multirow{30}{1em}{ \includegraphics[scale=0.35, trim = 20mm 60mm 120mm 40mm, clip]{fig3_legend.pdf}}\\ 
\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part2.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part2.pdf} &\\
\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part3.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part3.pdf} &\\
\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part4.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part4.pdf} & \\
\includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Post_part5.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 30mm 10mm 20mm, clip]{fig3_Cov_part5.pdf} &\\
\includegraphics[scale=0.25, trim = 0mm 5mm 10mm 20mm, clip]{fig3_Post_part6.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 5mm 10mm 20mm, clip]{fig3_Cov_part6.pdf} &
\end{tabular}
\end{center}
\caption{Mean posterior predictive functions (left) and covariance functions (right) of both the regular GP model and the HSGP model for different number of basis functions $m$ and for different values of the boundary factor $c$.}
  \label{fig3_Post_part1}
\end{figure}

@@

Additionally, there is a clear relation of the number of basis functions $m$ and the boundary factor $c$ with the lengthscale $\ell$ of the approximated function. Figure \ref{fig6_lscale_vs_J_vs_c} depicts how these three factors interact with each other in relation to 
% the performance of the HSGP approximation.
a close approximation of the HSGP model, in the case of a GP with square exponential covariance function and a single input dimension. (Gabi: Could it be good to build and include the same Figure for a Matern($\frac{3}{2}$)?). More precisely, for a given GP model (with a square exponential covariance function) with lengthscale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_lscale_vs_J_vs_c} shows the minimum $m$ required to achieve a close approximation in terms of satisfying Equation (\ref{diff_covs}). We have considered an approximation to be a close enough when the difference between densities of the approximate covariance function and the exact covariance function, $\varepsilon$ in Equation (\ref{diff_covs}), is below 1$\%$ of the density of the exact covariance function,
%
\begin{eqnarray*}
 \frac{\varepsilon}{\int k(\tau) \, d\tau} < 0.01.
\end{eqnarray*}

@@

As stated above, Figure \ref{fig6_lscale_vs_J_vs_c} provides the minimum lengthscale that can be closely approximated given $m$ and $c$. This information serves as a powerful diagnostic tool in determining if the obtained accuracy is acceptable. As the lenghscale $\ell$ controls the "wigglyness" of
the functional relationship, it strongly influences the difficulty of obtaining accurate
inference about the function from the data. Basically, if the lengthscale estimate is accurate, 
we can expect the HSGP approximation to be accurate as well. Thus, having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum lengthscale provided in Figure \ref{fig6_lscale_vs_J_vs_c}. If $\hat{\ell}$ exceeds this recommended minimum lengthscale, the approximation should be close enough. If, however, it does not exceed it, the approximation may be inaccurate and $m$ should be increased or $c$ decreased We may also use this diagnostic in a iterative procedure.
Starting from some initial guess of $\ell$, we can choose initial values for $m$ and $c$ and fit an HSGP model, then check the approximation accuracy, and, if not accurate enough because the estimated $\hat{\ell}$ is below the minimum lengthscale provided by Figure \ref{fig6_lscale_vs_J_vs_c}, repeat the process while increasing $m$ or decreasing $c$. Note that, as commented before, $c$ can not be decreased as much as desired because it is restricted to the lengthscale.
%replacing $\ell$ by the newly obtained estimate $\hat{\ell}$.

@@

@@

Figure \ref{fig6_lscale_vs_J_vs_c} was build for a GP with a unidimensional covariance function, and results in a surface depending on three variables, $m$, $c$ and $\ell$. An equivalent figure for a GP model with a two-dimensional covariance function would result in a surface depending on four variables, $m$, $c$, $\ell_1$ and $\ell_2$, which can not be directly represented. However, as an approximation we can use the unidimensional GP conclusions in Figure \ref{fig6_lscale_vs_J_vs_c} to analyze individually the different dimensions of a multidimensional GP model.

(Gabi: The following paragraph is another idea to introduce that the diagnosis plot in Figure \ref{fig6_lscale_vs_J_vs_c} depends on the kernel considered. At the moment we have built this figure only for the square exponential kernel.)

Ultimately, Figure \ref{fig6_lscale_vs_J_vs_c} showing the relationships among the number of basis functions, the boundary factor and the lengthscale of the function depends on the particular choice of covariance function. In this work we have built these relationships only for the square exponential covariance function. For other kernels, the Figure will be slightly different.

@@

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.4, trim = 10mm 95mm 5mm 30mm, clip]{fig7_legend.pdf}}\\
\vspace{-2mm}
\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part1.pdf}}
\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part2.pdf}}
\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part3.pdf}}\\
\vspace{-3mm}
\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part4.pdf}}
\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part5.pdf}}
\subfigure{\includegraphics[scale=0.29, trim = 10mm 31mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part6.pdf}}\\
\vspace{-3mm}
\subfigure{\includegraphics[scale=0.29, trim = 10mm 10mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part7.pdf}}
\subfigure{\includegraphics[scale=0.29, trim = 10mm 10mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part8.pdf}}
\subfigure{\includegraphics[scale=0.29, trim = 10mm 10mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part9.pdf}}\\
\vspace{-3mm}
\caption{Mean posterior predictions of both regular GP and HSGP models, fitted over various datasets drawn from square exponential GP models with different characteristic lengthscales ($lscale$) and same marginal variance ($\alpha$) as the data-generating functions (\textit{True function}).}
  \label{fig7_posterior_varing_lscale_part1}
\end{figure}

\begin{figure}
\begin{flushleft}
\begin{tabular}{ c c c c}
\includegraphics[scale=0.28, trim = 0mm 30mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part1.pdf} \vspace{0mm} & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 30mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part2.pdf} \vspace{0mm} & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 30mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part3.pdf} \vspace{0mm} & \hspace{-8mm} \multirow{3}{1em}{\vspace{-9mm} \includegraphics[scale=0.35, trim = 10mm 105mm 130mm 45mm, clip]{fig8_legend.pdf}}\\ 
\includegraphics[scale=0.28, trim = 0mm 10mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part4.pdf} \vspace{0mm}  & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 10mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part5.pdf} \vspace{0mm} & \hspace{-5mm} \includegraphics[scale=0.28, trim = 22mm 10mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part6.pdf} \vspace{0mm}
\end{tabular}
\end{flushleft}
\caption{Data-generating functional lengthscales (X-axis), of the various datasets illustrated in Figure \ref{fig7_posterior_varing_lscale_part1}, versus the corresponding lengthscale estimates from the regular GP and HSGP models (Y-axis). 95\% confident intervals of the lengthscale estimates are plotted as dot lines. The different plots represent the use of different number of basis functions $m$ in the HSGP model. The dashed black line represents the recommended minimum lengthscales provided by Figure \ref{fig6_lscale_vs_J_vs_c} that can be closely approximated by the HSGP model in every case.}
  \label{fig8_Tlscale_vs_Elscale}
\end{figure}

@@

This example consists of a simulated dataset with 250 ($i=1,\dots,n=250$) single draws of a Gaussian process prior with a Mat{\'e}rn($\nu$=3/2) covariance function and hyperparameters marginal variance $\alpha=1$ and lengthscale $\ell=0.15$, with corresponding inputs values $\bm{x}=\{x_i\}_{i=1}^{250}$ with $x_i \in [-1,1] \subset {\rm I\!R}$. To form the final noisy dataset $\bm{y}$, Gaussian noise $\sigma=0.2$ was added to the GP draws.

@@

\noindent where $I$ represents the identity matrix. The previous formulation corresponds to the latent form of a GP model, where $f(\bm{x})$ represents the underlying functional model to the noisy data, and it is modeled by a GP prior with a Mat{\'e}rn($\nu=3/2$) covariance function $k$. Saying that the function values $f(\bm{x})$ follows a GP model is equivalent to say that $f(\bm{x})$ are multivariate Gaussian distributed with covariance matrix $K$, where $K_{ij}=k(x_i,x_j,\theta)$.

@@

\begin{figure}[H]
\centering
\includegraphics[scale=0.50]{fig12_Posteriors_gaydata.pdf}
\caption{Posterior predictive distributions of the proposed HSGP model, the regular GP model, and the Splines model.}
  \label{fig12_Posteriors_gaydata}
\end{figure}

@@

Figure \ref{fig14_MSE_train_gaydata} shows the RMSE of the regular GP, HSGP and splines models, computed against the actual data, for training and test data, as a function of the number of basis functions $m$ and boundary factor $c$ for the HSGP model, and  knots for the splines model. We can see how the splines models does not extrapolate data properly. Figure \ref{fig14_1_time_gaydata} shows computational times, in seconds per iteration, as a function of the number of basis functions $m$ for the HSGP model and knots for the splines model. The computational times is represented in the Y-axis which is in a logarithmic scale. The HSGP model is on average roughly 10 times faster than the regular GP, in this particular case.

@@

\begin{figure}
\begin{center}
\begin{tabular}{ c c c}
\includegraphics[scale=0.35, trim = 0mm 33mm 35mm 20mm, clip]{fig15_truefun_exII.pdf} & \hspace{-5mm} \includegraphics[scale=0.35, trim = 20mm 33mm 35mm 20mm, clip]{fig15_gpfun_exII.pdf} & \hspace{-5mm}\multirow{-5.5}{*}{ \includegraphics[scale=0.35, trim = 150mm 20mm 10mm 10mm, clip]{fig15_colorbar_vertical_exII.pdf}}\\ 
\includegraphics[scale=0.35, trim = 0mm 22mm 35mm 20mm, clip]{fig15_bffun_exII.pdf}  & \hspace{-5mm} \includegraphics[scale=0.35, trim = 20mm 22mm 35mm 20mm, clip]{fig15_spfun_exII.pdf}
\end{tabular}
\end{center}
\caption{(a) Data-generating function. (b) Mean posterior predictive function of the GP model. (c) Mean posterior predictive function of the HSGP model. (d) Mean posterior predictive function of the splines (SP) model.}
  \label{fig15_Posteriors_exII}
\end{figure}

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.35, trim = 8mm 23mm 40mm 20mm, clip]{fig16_errorGP_exII.pdf}}
\subfigure{\includegraphics[scale=0.35, trim = 20mm 23mm 40mm 20mm, clip]{fig16_errorHSGP_exII.pdf}}
\subfigure{\includegraphics[scale=0.35, trim = 20mm 23mm 40mm 20mm, clip]{fig16_errorSP_exII.pdf}}
\subfigure{\includegraphics[scale=0.33, trim = 149mm 20mm 5mm 20mm, clip]{fig16_error_colorbar_exII.pdf}}
\caption{Mean error between the data-generating function and the GP (a), HSGP (b) and splines (c) models.}
  \label{fig16_errors_exII}
\end{figure}

@@

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig17_RMSE_exII.pdf}}
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig17_time_exII.pdf}}
\subfigure{\includegraphics[scale=0.40, trim = 22mm 45mm 120mm 0mm, clip]{fig17_legend_exII.pdf}}
\caption{Root mean square error (RMSE) (left) and computational time in seconds per iteration (right) of the different methods computed against the data-generating function, as a function of the boundary factor $c$, number of basis functions $m$ and knots.}
  \label{fig17_RMSE_exII}
\end{figure}

@@

Figure \ref{fig18_gpfun_diabetes} shows the mean posterior predictions of probabilities ($p_i$) of the three models, the regular GP, the HSGP, and the splines, fitted over the dataset with the 2 input dimensions ($D=2$) \textit{Glucose} and \textit{Pregnancy}. The binary observations $y_i$ are also plotted in the plots as colored points. For the HSGP model, $m_1=20$ and $m_2=20$ basis functions for each dimension, respectively, were used, which lead to a total of 400 multivariate basis functions. A boundary factor for each dimension $c_1=4$ and $c_2=4$ were used. For the splines model, 20 knots per dimension were used.
%
\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.35, trim = 8mm 23mm 40mm 20mm, clip]{fig18_gpfun_diabetes.pdf}}
\subfigure{\includegraphics[scale=0.35, trim = 20mm 23mm 40mm 20mm, clip]{fig18_bffun_diabetes.pdf}}
\subfigure{\includegraphics[scale=0.35, trim = 20mm 23mm 40mm 20mm, clip]{fig18_spfun_diabetes.pdf}}
\subfigure{\includegraphics[scale=0.31, trim = 149mm 18mm 5mm 20mm, clip]{fig18_legend_diabetes.pdf}}
\caption{(a) Mean posterior predictive functions of the GP model. (b) Mean posterior predictive functions of the HSGP model. (c) Mean posterior predictive functions of the splines (SP) model.}
  \label{fig18_gpfun_diabetes}
\end{figure}

>>>>

\usepackage{subfigure}
\usepackage{multirow,slashbox}
\usepackage{float}
\usepackage{soul}
\usepackage{xcolor}
\graphicspath{{./images/}}

@@

Ultimately, these recommendations lie on the relationships among the number of basis functions, the boundary factor and the lengthscale of the function, which depend on the particular choice of the kernel function. In this work we built these relationships for the square exponential covariance function and Matern ($\nu=3/2$) covariance function. For other kernels, the relationships will be slightly different, in function mainly of the smoothness or wigglyness of the kernel effects.

@@

%\begin{figure}
%\begin{center}
%\begin{tabular}{ c c c}
%\includegraphics[scale=0.25, trim = 0mm 19mm 10mm 20mm, clip]{fig3_Post_part1.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 19mm 10mm 20mm, clip]{fig3_Cov_part1.pdf} & \hspace{-6mm} \multirow{30}{1em}{ \includegraphics[scale=0.35, trim = 20mm 60mm 120mm 40mm, clip]{fig3_legend.pdf}}\\ 
%\includegraphics[scale=0.25, trim = 0mm 19mm 10mm 20mm, clip]{fig3_Post_part2.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 19mm 10mm 20mm, clip]{fig3_Cov_part2.pdf} &\\
%\includegraphics[scale=0.25, trim = 0mm 19mm 10mm 20mm, clip]{fig3_Post_part3.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 19mm 10mm 20mm, clip]{fig3_Cov_part3.pdf} &\\
%\includegraphics[scale=0.25, trim = 0mm 19mm 10mm 20mm, clip]{fig3_Post_part4.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 19mm 10mm 20mm, clip]{fig3_Cov_part4.pdf} & \\
%\includegraphics[scale=0.25, trim = 0mm 19mm 10mm 20mm, clip]{fig3_Post_part5.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 19mm 10mm 20mm, clip]{fig3_Cov_part5.pdf} &\\
%\includegraphics[scale=0.25, trim = 0mm 5mm 10mm 20mm, clip]{fig3_Post_part6.pdf} & \hspace{-5mm} \includegraphics[scale=0.25, trim = 0mm 5mm 10mm 20mm, clip]{fig3_Cov_part6.pdf} &
%\end{tabular}
%\end{center}
%\caption{Mean posterior predictive functions (left) and covariance functions (right) of both the regular GP model and the HSGP model for different number of basis functions $m$ and for different values of the boundary factor $c$.}
%  \label{fig3_Post_part1}
%\end{figure}

\begin{figure}
\begin{tabular}{ c c c c c}
\includegraphics[scale=0.20, trim = 10mm 24mm 10mm 20mm, clip]{fig3_Post_part1.pdf} & \hspace{-5mm} 
\includegraphics[scale=0.20, trim = 5mm 24mm 10mm 20mm, clip]{fig3_Cov_part1.pdf} &  \hspace{-0mm}
\includegraphics[scale=0.20, trim = 5mm 24mm 10mm 20mm, clip]{fig3_Post_part4.pdf} & \hspace{-5mm} 
\includegraphics[scale=0.20, trim = 5mm 24mm 10mm 20mm, clip]{fig3_Cov_part4.pdf} & \\ 
\includegraphics[scale=0.20, trim = 10mm 24mm 10mm 20mm, clip]{fig3_Post_part2.pdf} & \hspace{-5mm} 
\includegraphics[scale=0.20, trim = 5mm 24mm 10mm 20mm, clip]{fig3_Cov_part2.pdf} & \hspace{-0mm}
\includegraphics[scale=0.20, trim = 5mm 24mm 10mm 20mm, clip]{fig3_Post_part5.pdf} & \hspace{-5mm} 
\includegraphics[scale=0.20, trim = 5mm 24mm 10mm 20mm, clip]{fig3_Cov_part5.pdf} & \hspace{-5mm} 
\includegraphics[scale=0.35, trim = 25mm 63mm 120mm 50mm, clip]{fig3_legend.pdf}\\
\includegraphics[scale=0.20, trim = 10mm 5mm 10mm 20mm, clip]{fig3_Post_part3.pdf} & \hspace{-5mm} 
\includegraphics[scale=0.20, trim = 5mm 5mm 10mm 20mm, clip]{fig3_Cov_part3.pdf} & \hspace{-0mm}
\includegraphics[scale=0.20, trim = 5mm 5mm 10mm 20mm, clip]{fig3_Post_part6.pdf} & \hspace{-5mm} 
\includegraphics[scale=0.20, trim = 5mm 5mm 10mm 20mm, clip]{fig3_Cov_part6.pdf} & 
\end{tabular}
\caption{Mean posterior predictive functions (left-first and right-first columns) and covariance functions (left-second and right-second columns) of both the regular GP model and the HSGP model for different number of basis functions $m$ and for different values of the boundary factor $c$.}
  \label{fig3_Post_part1}
\end{figure}

@@

Additionally, there is a clear relation of the number of basis functions $m$ and the boundary factor $c$ with the lengthscale $\ell$ of the approximated function. Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} depicts how these three factors interact with each other in relation to 
% the performance of the HSGP approximation.
a close approximation of the HSGP model, in the cases of a GP with square exponential covariance function and Mat\'ern ($\nu=3/2$) covariance function, respectively, and a single input dimension. More precisely, for a given GP model (with a square exponential covariance function) with lengthscale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_lscale_vs_J_vs_c} shows the minimum $m$ required to achieve a close approximation in terms of satisfying Equation (\ref{diff_covs}). Similarly for Figure \ref{fig6_lscale_vs_J_vs_c_Matern} in the case of a Mat\'ern ($\nu=3/2$) covariance function. We have considered an approximation to be a close enough when the difference between densities of the approximate covariance function and the exact covariance function, $\varepsilon$ in Equation (\ref{diff_covs}), is below 1$\%$ of the density of the exact covariance function,
%
\begin{eqnarray*}
 \frac{\varepsilon}{\int k(\tau) \, d\tau} < 0.01.
\end{eqnarray*}

@@

As stated above, Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} provide the minimum lengthscale that can be closely approximated given $m$ and $c$. This information serves as a powerful diagnostic tool in determining if the obtained accuracy is acceptable. As the lenghscale $\ell$ controls the "wigglyness" of
the functional relationship, it strongly influences the difficulty of obtaining accurate
inference about the function from the data. Basically, if the lengthscale estimate is accurate, 
we can expect the HSGP approximation to be accurate as well. Thus, having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum lengthscale provided in Figure \ref{fig6_lscale_vs_J_vs_c}. If $\hat{\ell}$ exceeds this recommended minimum lengthscale, the approximation should be close enough. If, however, it does not exceed it, the approximation may be inaccurate and $m$ should be increased or $c$ decreased. We may also use this diagnostic in a iterative procedure.
Starting from some initial guess of $\ell$, we can choose initial values for $m$ and $c$ and fit an HSGP model, then check the approximation accuracy, and, if not accurate enough because the estimated $\hat{\ell}$ is below the minimum lengthscale provided by Figure \ref{fig6_lscale_vs_J_vs_c}, repeat the process while increasing $m$ or decreasing $c$. Note that, as commented before, $c$ can not be decreased as much as desired because it is restricted to the lengthscale.
%replacing $\ell$ by the newly obtained estimate $\hat{\ell}$.

@@

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig6_lscale_vs_J_vs_c_Matern.pdf}}
\hspace{3mm}
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig6_lscale_vs_J_vs_c_zoomin_Matern.pdf}}
\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ ($c = \frac{L}{S}$) and the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$), in the case of a Mat\`ern($\frac{3}{2}$) covariance function. The right-side plot is a zoom in of the left-side plot.}
  \label{fig6_lscale_vs_J_vs_c_Matern}
\end{figure}

@@

Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} were build for a GP with a unidimensional covariance function, and result in a surface depending on three variables, $m$, $c$ and $\ell$. An equivalent figure for a GP model with a two-dimensional covariance function would result in a surface depending on four variables, $m$, $c$, $\ell_1$ and $\ell_2$, which can not be directly represented. However, as an approximation we can use the unidimensional GP conclusions in Figure \ref{fig6_lscale_vs_J_vs_c} to analyze individually the different dimensions of a multidimensional GP model.
\begin{center}
•
\end{center}

@@

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.4, trim = 10mm 100mm 5mm 30mm, clip]{fig7_legend.pdf}}\\
\vspace{-2.5mm}
\subfigure{\includegraphics[scale=0.29, trim = 10mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part1.pdf}}
\subfigure{\includegraphics[scale=0.29, trim = 10mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part2.pdf}}
\subfigure{\includegraphics[scale=0.29, trim = 10mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part3.pdf}}\\
\vspace{-3mm}
\subfigure{\includegraphics[scale=0.29, trim = 10mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part4.pdf}}
\subfigure{\includegraphics[scale=0.29, trim = 10mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part5.pdf}}
\subfigure{\includegraphics[scale=0.29, trim = 10mm 25.5mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part6.pdf}}\\
\vspace{-3mm}
\subfigure{\includegraphics[scale=0.29, trim = 10mm 7mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part7.pdf}}
\subfigure{\includegraphics[scale=0.29, trim = 10mm 7mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part8.pdf}}
\subfigure{\includegraphics[scale=0.29, trim = 10mm 7mm 10mm 19mm, clip]{fig7_posterior_varing_lscale_part9.pdf}}\\
\vspace{-3mm}
\caption{Mean posterior predictions of both regular GP and HSGP models, fitted over various datasets drawn from square exponential GP models with different characteristic lengthscales ($lscale$) and same marginal variance ($\alpha$) as the data-generating functions (\textit{True function}).}
  \label{fig7_posterior_varing_lscale_part1}
\end{figure}

\begin{figure}
\begin{flushleft}
\begin{tabular}{ccc}
\multicolumn{3}{c}{ \includegraphics[scale=0.35, trim = 10mm 115mm 80mm 45mm, clip]{fig8_legend.pdf}}\\
\includegraphics[scale=0.28, trim = 0mm 26mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part1.pdf} & \hspace{-4mm}
\includegraphics[scale=0.28, trim = 20mm 26mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part2.pdf} & \hspace{-4mm} \includegraphics[scale=0.28, trim = 20mm 26mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part3.pdf} \\ 
\includegraphics[scale=0.28, trim = 0mm 5mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part4.pdf} & \hspace{-4mm} \includegraphics[scale=0.28, trim = 20mm 5mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part5.pdf} & \hspace{-4mm} \includegraphics[scale=0.28, trim = 20mm 5mm 10mm 20mm, clip]{fig8_Tlscale_vs_Elscale_part6.pdf}
\end{tabular}
\end{flushleft}
\caption{Data-generating functional lengthscales (X-axis), of the various datasets illustrated in Figure \ref{fig7_posterior_varing_lscale_part1}, versus the corresponding lengthscale estimates from the regular GP and HSGP models (Y-axis). 95\% confident intervals of the lengthscale estimates are plotted as dot lines. The different plots represent the use of different number of basis functions $m$ in the HSGP model. The dashed black line represents the recommended minimum lengthscales provided by Figure \ref{fig6_lscale_vs_J_vs_c} that can be closely approximated by the HSGP model in every case.}
  \label{fig8_Tlscale_vs_Elscale}
\end{figure}

@@

This example consists of a simulated dataset with 250 ($i=1,\dots,n=250$) single draws from a Gaussian process prior with a Mat{\'e}rn($\nu$=3/2) covariance function and hyperparameters marginal variance $\alpha=1$ and lengthscale $\ell=0.15$, with corresponding inputs values $\bm{x}=\{x_i\}_{i=1}^{250}$ with $x_i \in [-1,1] \subset {\rm I\!R}$. To form the final noisy dataset $\bm{y}$, Gaussian noise $\sigma=0.2$ was added to the GP draws.

@@

\noindent where $I$ represents the identity matrix. The previous formulation corresponds to the latent form of a GP model, where $f(\bm{x})$ represents the underlying functional model to the noisy data and is modeled by a GP prior with a Mat{\'e}rn($\nu=3/2$) covariance function $k$. Saying that the function values $f(\bm{x})$ follows a GP model is equivalent to say that $f(\bm{x})$ are multivariate Gaussian distributed with covariance matrix $K$, where $K_{ij}=k(x_i,x_j,\theta)$.

@@

\begin{figure}
\centering
\includegraphics[scale=0.50]{fig12_Posteriors_gaydata.pdf}
\caption{Posterior predictive distributions of the proposed HSGP model, the regular GP model, and the Splines model.}
  \label{fig12_Posteriors_gaydata}
\end{figure}

@@

Figure \ref{fig14_MSE_train_gaydata} shows the RMSE of the regular GP, HSGP and splines models, computed against the actual data, for training and test data, as a function of the number of basis functions $m$ and boundary factor $c$ for the HSGP model, and  knots for the splines model. We can see how the splines models does not extrapolate data properly.

Figure \ref{fig14_1_time_gaydata} shows computational times, in seconds per iteration, as a function of the number of basis functions $m$ for the HSGP model and knots for the splines model. The computational times is represented in the Y-axis which is in a logarithmic scale. The HSGP model is on average roughly 10 times faster than the regular GP, in this particular case.

@@

\begin{figure}
\begin{center}
\begin{tabular}{ c c c}
\includegraphics[scale=0.35, trim = 0mm 33mm 35mm 20mm, clip]{fig15_truefun_exII.pdf} & \hspace{-5mm} \includegraphics[scale=0.35, trim = 19mm 33mm 35mm 20mm, clip]{fig15_gpfun_exII.pdf} & \hspace{-5mm}\multirow{-5.5}{*}{ \includegraphics[scale=0.35, trim = 150mm 20mm 0mm 10mm, clip]{fig15_colorbar_vertical_exII.pdf}}\\ 
\includegraphics[scale=0.35, trim = 0mm 22mm 35mm 20mm, clip]{fig15_bffun_exII.pdf}  & \hspace{-5mm} \includegraphics[scale=0.35, trim = 19mm 22mm 35mm 20mm, clip]{fig15_spfun_exII.pdf}
\end{tabular}
\end{center}
\caption{(a) Data-generating function. (b) Mean posterior predictive function of the GP model. (c) Mean posterior predictive function of the HSGP model. (d) Mean posterior predictive function of the splines (SP) model.}
  \label{fig15_Posteriors_exII}
\end{figure}

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.35, trim = 8mm 23mm 39.2mm 20mm, clip]{fig16_errorGP_exII.pdf}}
\subfigure{\includegraphics[scale=0.35, trim = 19mm 23mm 39.2mm 20mm, clip]{fig16_errorHSGP_exII.pdf}}
\subfigure{\includegraphics[scale=0.35, trim = 19mm 23mm 39.2mm 20mm, clip]{fig16_errorSP_exII.pdf}}
\subfigure{\includegraphics[scale=0.33, trim = 149mm 20mm 0mm 20mm, clip]{fig16_error_colorbar_exII.pdf}}
\caption{Mean error between the data-generating function and the GP (a), HSGP (b) and splines (c) models.}
  \label{fig16_errors_exII}
\end{figure}

@@

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig17_RMSE_exII.pdf}} \hspace{2mm}
\subfigure{\includegraphics[scale=0.35, trim = 0mm 0mm 10mm 0mm, clip]{fig17_time_exII.pdf}}
\subfigure{\includegraphics[scale=0.40, trim = 22mm 45mm 120mm 0mm, clip]{fig17_legend_exII.pdf}}
\caption{Root mean square error (RMSE) (left) and computational time in seconds per iteration (right) of the different methods computed against the data-generating function, as a function of the boundary factor $c$, number of basis functions $m$ and knots.}
  \label{fig17_RMSE_exII}
\end{figure}

@@

Figure \ref{fig18_gpfun_diabetes} shows the mean posterior predictions of probabilities ($p_i$) of the three models, the regular GP, the HSGP, and the splines, fitted over the dataset with the 2 input dimensions ($D=2$) \textit{Glucose} and \textit{Pregnancy}. The binary observations $y_i$ are also plotted in the plots as colored points. For the HSGP model, $m_1=20$ and $m_2=20$ basis functions for each dimension, respectively, were used, which lead to a total of 400 multivariate basis functions. A boundary factor for each dimension $c_1=4$ and $c_2=4$ were used. For the splines model, 20 knots per dimension were used.
%
\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.35, trim = 12mm 23mm 35mm 20mm, clip]{fig18_gpfun_diabetes.pdf}}
\subfigure{\includegraphics[scale=0.35, trim = 20mm 23mm 35mm 20mm, clip]{fig18_bffun_diabetes.pdf}}
\subfigure{\includegraphics[scale=0.35, trim = 20mm 23mm 38mm 20mm, clip]{fig18_spfun_diabetes.pdf}}
\subfigure{\includegraphics[scale=0.31, trim = 150mm 13mm 2mm 15mm, clip]{fig18_legend_diabetes.pdf}}
\caption{(a) Mean posterior predictive functions of the GP model. (b) Mean posterior predictive functions of the HSGP model. (c) Mean posterior predictive functions of the splines (SP) model.}
  \label{fig18_gpfun_diabetes}
\end{figure}
